---
title: "Vignette for JiangZhihangTools"
author: "Zhihang Jiang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

#Info
  This is package is built for quiz3 and homework3, which inclde all functions in pervious homeworks and quizs. It includes three data set and 14 functions. The package can be used to calculate mean, variance and standard deviation. Also it can calcuat MLE for most of distribution and there is an efficient way for gamma distribution. Then it is very useful for matrix calculation. 
  
#data
there are 3 data set in the package
```{r}
data2007 <- read.csv("http://users.stat.umn.edu/~almquist/3811_examples/gapminder2007ex.csv")
d <- read.table(url("http://www.stat.umn.edu/geyer/3701/data/q1p4.txt"),header = TRUE)
m<-url("http://www.stat.umn.edu/geyer/3701/data/q2p3.rda")
```


#functions

##function1
it can calculate Mean, Variance, SD
```{r}
func1 <- function(x){
  a = sum(x)/length(x)
  b = sum((x-a)^2)/length(x)
  c = sqrt(b)
  return(list(mean=a,var=b,sd=c))
}
func1(rnorm(10))
```

##function2
It also calculate Mean,Variance,SD and it can check the input data
```{r}
func2 <- function(x){
  stopifnot(is.numeric(x))
  stopifnot(length(x)!=0)
  stopifnot(is.finite(x))
  stopifnot(!is.na(x))
  stopifnot(!is.nan(x))
  
  a = sum(x)/length(x)
  b = sum((x-a)^2)/length(x)
  c = sqrt(b)
  return(list(mean=a,var=b,sd=c))
}
func2(rnorm(10))
```

##function3
This function3 computes the liklihood of a gamma distribution
```{r}
func3 <- function(x){
  alpha <- pi
  log <- function(alpha)
    sum(dgamma(x, shape = alpha, log = TRUE))
  interval <- mean(x) + c(-1,1) * 3 * sd(x)
  interval <- pmax(mean(x) / 1e3, interval)
  
  oout<- optimize(log, maximum = TRUE, interval)
  return (oout$maximum)
}
x <- scan(url("http://www.stat.umn.edu/geyer/3701/data/q1p3.txt"))
func3(x)
```


##function4
function4 calculates weigthed Mean,Variance,SD
```{r}
func4 <- function(d){
  
  a = sum(d$x * d$p)
  b = sum(((d$x - a)^2) * d$p)
  c = sqrt(b)
  return(list(mean=a,var=b,sd=c))
  
}

func4(d)
```


##function5
function5 calculates Mean,Variance,SD with checking the input data
```{r}
func5 <- function(d){
  
  stopifnot(is.numeric(d$x))
  stopifnot(is.numeric(d$p))
  
  stopifnot(length(d$x)!=0)
  stopifnot(length(d$p)!=0)
  
  stopifnot(is.finite(d$x))
  stopifnot(is.finite(d$p))
  
  stopifnot(!is.na(d$x))
  stopifnot(!is.na(d$p))
  
  stopifnot(!is.nan(d$x))
  stopifnot(!is.nan(d$p))
  
  stopifnot(all.equal(sum(d$p),1))
  
  a = sum(d$x * d$p)
  b = sum(((d$x - a)^2) * d$p)
  c = sqrt(b)
  return(list(mean=a,var=b,sd=c))
  
}
a<-read.table(url("http://www.stat.umn.edu/geyer/3701/data/q1p4.txt"),
    header = TRUE)
func5(a)
```

##function6
function6 calculates weighted Mean,Variance,SD with checking the input data
```{r}
func6<-function(x,y)
{
  if(!length(y)>0){
    print("length of the second input is not positive")
  }else if(!length(x)>0){
    print("the length of first input is not positive")
  }else if(!is.numeric(x)){
    print("the first input is not numeric")
  }else if(!is.numeric(y)){
    print("the second input is not numeric")
  }else if(sum(is.finite(x))!=length(x)){
    print("the first input has infinite elements")
  }else if(sum(is.finite(y))!=length(y)){
    print("the second input has infinite elements")
  }else if(length(y)!=length(x)){
    print("the lenghs of two input are not equal")
  }else if(!isTRUE(all.equal(sum(y),1))){
    print("the sum of second input is not equal to 1")
  }else if(sum(y>0)!=length(y)){
    print("second input have negative elements")
  }else{
    mu<-sum((x)*y)
    si<-sum(((x-mu)^2)*y)
    data.frame(mean = mu,variance = si, sd = sqrt(si))
  }
}
func6(1,Inf)
func6(1,"a")
func6(double(0),1)
func6(c(1,2,3),c(0,2,1))
```

##function7
Computes the liklihood of a given distribution for data x
```{r}
func7 <- function(x, func, interval){
  
  f7 <- function(theta, x)
  {sum(func(theta, x))}
  
  oout<- optimize(f7, maximum = TRUE, interval, x=x)
  return(oout$maximum)
} 
x1=rgamma(100,3)
f = function(theta, x) dgamma(x, shape = theta, log = TRUE)
func7(x1,f,c(0,3))
```

##function8
function8 calculates matrix multiplcation $x^T A^{-1} x$
```{r}
func8<-function(a,x){
  stopifnot(is.matrix(a))
  stopifnot(is.vector(x))
  stopifnot(is.numeric(x))
  stopifnot(is.numeric(a))
  stopifnot(is.finite(x))
  stopifnot(is.finite(a))
  stopifnot(nrow(a)==ncol(a))
  stopifnot(nrow(a)==length(x))
  m=solve(a,x)
  sum(x*m)
}
load(url("http://www.stat.umn.edu/geyer/3701/data/q2p1.rda"))
func8(a,x)
```

##function9
function9 calculates matrix multiplcation $x^T A^{-1} x$ as a binary operator
```{r}
"%func9%" <-function(a,x){
  stopifnot(is.matrix(a))
  stopifnot(is.vector(x))
  stopifnot(is.numeric(x))
  stopifnot(is.numeric(a))
  stopifnot(is.finite(x))
  stopifnot(is.finite(a))
  stopifnot(nrow(a)==ncol(a))
  stopifnot(nrow(a)==length(x))
  m=solve(a,x)
  sum(x*m)
}
load(url("http://www.stat.umn.edu/geyer/3701/data/q2p1.rda"))
a%func9%x
```

##function10
function10 standardizes matrix
```{r}
func10<-function(x){
  stopifnot(nrow(x)>1)
  stopifnot(is.matrix(x))
  stopifnot(is.numeric(x))
  stopifnot(is.finite(x))
  for(i in 1:ncol(x)){
    x[,i]=(x[,i]-mean(x[,i]))/sd(x[,i])
  }
  x
}
load(url("http://www.stat.umn.edu/geyer/3701/data/q2p3.rda"))
func10(a)
```
##function11
function11 standardizes matrix without loop
```{r}
func11<-function(x){
  stopifnot(is.matrix(x))
  stopifnot(is.numeric(x))
  stopifnot(is.finite(x))
  FUN<-function(x)
  {
    x<-(x-mean(x))/sd(x)
  }
  apply(x,2,FUN)    
}
load(url("http://www.stat.umn.edu/geyer/3701/data/q2p3.rda"))
func11(a)
```

##myapply
It is similar with apply function
```{r}
myapply <- function(x,MARGIN,FUN,...)
{
  if(length(dim(x))!=2)
  {
    stop("matrix is not 2d")
  }
  if(!(MARGIN %in% c(1,2)))
  {
    stop("margin is not in 1 or 2")
  }
  R=dim(x)[1]
  C=dim(x)[2]
  f= match.fun(FUN)
  if(MARGIN==1)
  {
    result=list()
    for(i in 1:R)
    {
      result[[i]]=f(x[i,],...)
    }
  }else if(MARGIN==2)
  {
    result=list()
    for(j in 1:C)
    {
      result[[j]]=f(x[,j],...)
    }
  }
  return(simplify2array(result))
}
fred=matrix(1:6,nrow = 3)
rownames(fred)=c("red","white","blue")
colnames(fred)=c("car","truck")
myapply(fred, 1, mean)
```

##MLEfunc
It is same as function7
```{r}
MLEfunc<-function(x,f,interval){
  logl<-function(theta){
  sum(f(theta,x))
  }
  oout<-optimize(logl,maximum=TRUE,interval)
  oout$maximum
}
x1=rgamma(100,3)
f = function(theta, x) dgamma(x, shape = theta, log = TRUE)
MLEfunc(x1,f,c(0,3))
```

##plot function
Wrapper function for ggplot2 for data d
```{r}
plotMyData<-function(x){
  library(magrittr)
  x%>% ggplot2::ggplot()+ggplot2::aes(x=x, y=p)+ggplot2::geom_point()
}
plotMyData(d)
```

##select function
```{r}
selectmydata<-function(x){
  library(magrittr)
  library(tidyverse)
  library(dplyr)
  xa<-data2007%>%
select(gdpPercap, pop) %>% 
mutate(gdp = pop * gdpPercap)
  return(xa)
}
selectmydata(data2007)
```

#more example
It is for making up Quiz 3
```{r}
ggplot2::ggplot(d)+ggplot2::aes(x=x, y=p)+ggplot2::geom_point()
ggplot2::ggplot(d)+ggplot2::aes(x=x, y=p)+ggplot2::geom_line()
ggplot2::ggplot(d)+ggplot2::aes(x=x, y=p)+ggplot2::geom_area()
```
